\stktokens{}: Enforcing Well-bracketed Control Flow and Stack Encapsulation using Linear Capabilities

### 2018-07-06 fully-abstract-comp/popl2019/paper.tex
Typo
Typo.

% split/splice
Instructions for reducing the authority of capabilities are commonplace on capability machines as they allow us to limit what a capability can do before it is passed away.
For normal capabilities, reduction of authority can be done without actually giving up any authority if the capability is duplicated first.
With linear capabilities authority cannot be preserved in this fashion as they are non-duplicable.
In order to make a lossless reduction of the range of authority, \trgcm{} provides special hardware support in the form of \texttt{split} and \texttt{splice}.
The \texttt{split} instruction takes a capability with range of authority $[\var{base},\var{end}]$ and an address $n$ and creates two new capabilities one with $[\var{base},n]$ as its range of authority and the other with $[n+1,\var{end}]$ as it range of authority.
Everything else, i.e.\ permission, linearity and current address, is copied without change to the new capabilities.
With \texttt{split} we can reduce the range of authority of a linear capability without losing any authority as we retain it in the other capability.
However, if we only have \texttt{split}, then in some sense splitting a capability is not completely lossless as we may have authority over the same range of addresses, but we have that authority in two capabilities rather than just one.
In order to make up for this, we have a \texttt{splice} instruction that essentially does the inverse of \texttt{split}.
Namely given two capabilities with adjacent ranges of authority and the same permissions and linearity, it splices them together into one capability.
The two instructions work in the same way for seal sets.
We do not provide special support for lossless reduction of capability permissions, but this could probably be achieved with more fine-grained permissions.
This would also allow linear capabilities to have aliases, but only by linear capabilities with disjointed permissions.

At this time, the executable configuration is our only way to specify the state of the capability machine.
We would, however, much prefer a notion of a program that only describe the essential parts of an entity on the machine and only later instrument the program with whatever necessary to make it into an executable configuration.
In later sections, we describe different entities on the capability machine that interact with each other.
In order for them to interact, we need to be able to link them together and to this end, we introduce a notion of a component.
A component is basically a program with open ends in the form of imports.
A component also have exports allowing other programs to link with them.
We define the notion of a component in Figure~\ref{fig:target-component-and-linking}.
% Structure
A base component $\comp_0$ consists of the following: a code memory segment, a data memory segment, a list of imported symbols, a list of exported symbols, two lists specifying the available seals seals\footnote{We will return to the seals in later sections.}, and a set of all the linear addresses.
The import list specifies where in memory imports should be placed, and imports are matched to exports via their symbols.
The exports are words associated with a symbol.
%base = library
A base component can be seen as library implementations because it does not specify a main entry point of execution.
Components are either a library component or a base component with a main in the form of a pair of sealed capabilities which can be seen as a program that still needs to be linked with libraries.
% Linking .
Components are combined into new components by linking them together.
Libraries can be linked together and incomplete programs can be linked with libraries, but two incomplete programs cannot be linked with each other.
Two components can be linked when their memories, seals, and linear addresses disjointed.
They are combined by taking the union of each of their constituents.
For every import that is satisfied by an export of the other component, in the sense that they have the same symbol, the data memory is updated to have the exported word on the imported address.
The satisfied imports are removed from the import list in the resulting linked component and the exports are updated to be the exports of the two components.
In Figure~\ref{fig:target-component-and-linking}, $\bowtie$ defines linking.
\todo[inline]{Add better separation between component and linking in figure.}

>>>>

% split/splice
Instructions for reducing the authority of capabilities are commonplace on capability machines as they allow us to limit what a capability can do before it is passed away.
For normal capabilities, reduction of authority can be done without actually giving up any authority if the capability is duplicated first.
With linear capabilities authority cannot be preserved in this fashion as they are non-duplicable.
In order to make a lossless reduction of the range of authority, \trgcm{} provides special hardware support in the form of \texttt{split} and \texttt{splice}.
The \texttt{split} instruction takes a capability with range of authority $[\var{base},\var{end}]$ and an address $n$ and creates two new capabilities one with $[\var{base},n]$ as its range of authority and the other with $[n+1,\var{end}]$ as it range of authority.
Everything else, i.e.\ permission, linearity and current address, is copied without change to the new capabilities.
With \texttt{split} we can reduce the range of authority of a linear capability without losing any authority as we retain it in the other capability.
However, if we only have \texttt{split}, then in some sense splitting a capability is not completely lossless as we may have authority over the same range of addresses, but we have that authority in two capabilities rather than just one.
In order to make up for this, we have a \texttt{splice} instruction that essentially does the inverse of \texttt{split}.
Namely given two capabilities with adjacent ranges of authority and the same permissions and linearity, it splices them together into one capability.
The two instructions work in the same way for seal sets.
We do not provide special support for lossless reduction of capability permissions, but this could probably be achieved with more fine-grained permissions.
This would also allow linear capabilities to have aliases, but only by linear capabilities with disjoint permissions.

At this time, the executable configuration is our only way to specify the state of the capability machine.
We would, however, much prefer a notion of a program that only describe the essential parts of an entity on the machine and only later instrument the program with whatever necessary to make it into an executable configuration.
In later sections, we describe different entities on the capability machine that interact with each other.
In order for them to interact, we need to be able to link them together and to this end, we introduce a notion of a component.
A component is basically a program with open ends in the form of imports.
A component also has exports allowing other programs to link with it.
We define the notion of a component in Figure~\ref{fig:target-component-and-linking}.
% Structure
A base component $\comp_0$ consists of the following: a code memory segment, a data memory segment, a list of imported symbols, a list of exported symbols, two lists specifying the available seals seals\footnote{We will return to the seals in later sections.}, and a set of all the linear addresses.
The import list specifies where in memory imports should be placed, and imports are matched to exports via their symbols.
The exports are words associated with a symbol.
%base = library
A base component can be seen as library implementations because it does not specify a main entry point of execution.
Components are either a library component or a base component with a main in the form of a pair of sealed capabilities which can be seen as a program that still needs to be linked with libraries.
% Linking .
Components are combined into new components by linking them together.
Libraries can be linked together and incomplete programs can be linked with libraries, but two incomplete programs cannot be linked with each other.
Two components can be linked when their memories, seals, and linear addresses disjointed.
They are combined by taking the union of each of their constituents.
For every import that is satisfied by an export of the other component, in the sense that they have the same symbol, the data memory is updated to have the exported word on the imported address.
The satisfied imports are removed from the import list in the resulting linked component and the exports are updated to be the exports of the two components.
In Figure~\ref{fig:target-component-and-linking}, $\bowtie$ defines linking.
\todo[inline]{Add better separation between component and linking in figure.}
